---
title: "Structural_Bio"
author: "Brandon Gonzalez"
date: "February 14, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Read in the data
```{r}
protein_data <- read.csv(header=TRUE,file = "Data_Export_Summary.csv")
protein_df <- as.matrix(protein_data[2:6])
rownames(protein_df) <- protein_data$Experimental.Method
                        
```

#Lets determine the percentage of structures solved by X-Ray and Electron Microscopy
```{r}
protein_amounts <- protein_df[ ,"Proteins" ]
percents <- round(protein_df[,"Total"]/ sum(protein_df[,"Total"]),3)
names(percents) <- rownames(protein_df)
percents
percent_protein = round((sum(protein_df[,"Proteins"])/sum(protein_data$Total)),2)
barplot(height = protein_amounts, names.arg = rownames(protein_amounts), main = "Proteins", xlab = percent_protein)


```
The data base is 93% proteins with most of the proteins elucidated from X-Ray crystallography

##Bio3d
```{r}
library(bio3d)

```
#lets read HSG1 pdb, print out a summary
A pdb file contains atoms with their xyz coordinates at the most basic sense.
```{r}
hsg_pdb <- read.pdb("1hsg")
print(hsg_pdb)
```
We can see from the summary that there are 1514 atoms in the protein with a total of 198 residues, 172 atoms in non-protein,2 non protein reisdues: 1 MK1(merk1 drug) and 127 waters (HOH).
```{r}
#this wil print out the atom col, but it also tells us that its a dataframe
#str(pdb$atom) 

#print summary of coordinate data
hsg_pdb$xyz
```

```{r}
#select only protein atoms, and then only mk1. Then write as new pdb files
p.inds <- atom.select(hsg_pdb,"protein")
mk1.inds <- atom.select(hsg_pdb,"ligand")

p.trimmed <- trim(hsg_pdb, inds=p.inds)
mk1.trimmed <- trim(hsg_pdb, inds=mk1.inds)

  
write.pdb(pdb=p.trimmed, file = "Protein_only.pdb")
write.pdb(pdb=mk1.trimmed, file = "Mk1_only.pdb")
```


