---
title: "Find_A_Gene"
author: "Brandon Gonzalez"
date: "February 24, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(bio3d)
filename = "data/muscle_alignment"
```

```{r}
alignment = read.fasta(filename)

```


```{r}
plot.fasta(alignment, hc=TRUE)
```

```{r}
seqMatrix <- seqidentity(alignment)
heatmap(seqMatrix)
```

```{r}
heatmap(seqMatrix, cexRow = .7, cexCol = .5, margins = c(8,2))
#par(mai=c(2,2,0,2),pin=c(1,4.4))

```


Lets get a consensus from the alignment, the default cutoff identity value is 0.6.
With this consensus sequence I will then query pdb to find protein structures that are similar. 
```{r}
con <- consensus(alignment)
print(con$seq[1:50])
```

```{r}
pdbs <- blast.pdb(seq=con$seq)

```
Lets see the first 5 hits.
```{r}
pdbs$hit.tbl[1:5,]
#lets get the first 3 hits and get look them up on PDB database
selectedPdbs <- pdbs$hit.tbl[1:3,]
ids <- selectedPdbs$pdb.id
```

```{r}
annotated <- pdb.annotate(ids, unique = TRUE)
annotated
```
```{r}
pdb_df <- data.frame(annotated$experimentalTechnique,annotated$resolution, annotated$source, selectedPdbs$evalue, selectedPdbs$identity)
#pdb_df <- rownames(annotated$structureId)
pdb_df
```

Based on these results, I am a little  uncomfortable using these proteins because they arent as from a family as related as I want. I will do the heat map, consensus, and PDB query again.

```{r}
filename = "data/muscle_alignmentNoH"

alignment2 <- read.fasta(filename)
mtrix <- seqidentity(alignment2)
heatmap(mtrix, cexRow = .7, cexCol = .5, margins = c(8,2))
```
```{r}
con <- consensus(alignment2)
print(con$seq[1:50])
```

```{r}
#look up PDBS
pdbs <- blast.pdb(seq=con$seq)
#only want the first three again
selectedPdbs <- pdbs$hit.tbl[1:3,]
ids <- selectedPdbs$pdb.id

annotated <- pdb.annotate(ids, unique = TRUE)

```

### Looks like they are the same proteins, thats good then, these results mean theyre independent of methodology and that give me much more confidence in them.

```{r}
pdb_df <- data.frame(annotated$experimentalTechnique,annotated$resolution, annotated$source, selectedPdbs$evalue, selectedPdbs$identity)
row.names(pdb_df) <- ids
pdb_df
```
To visualize in VMD, download the PDBs locally and open them up.

```{r}
#Download pdb files locally

#get.pdb(ids)
```

```{r}

```

