---
title: "RNA Seq Analysis of Hines et al."
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Importing countData and colData
```{r}

counts <- read.csv("data/airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <- read.csv("data/airway_metadata.csv", stringsAsFactors = FALSE)

```
## Preview of data
```{r}
head(counts)

```
```{r}
head(metadata)
```

### The colnames of the counts data must match the ids of the metadata for DESeq2 to work 
```{r}
#the colnames must mat
colnames(counts)
```

```{r}
#all() checks that everything returned is TRUE or else it will return false, lets use this to check that the Ids match
all( colnames(counts)[-1] == metadata$id )

```

### Lets find out which columns in *counts* are the control conditions.
Then we find out what is the mean of count value for each gene.

```{r}
cntrl.inds = metadata$dex == "control"
#get the ids of the cntrl cells
cntrl.ids <- metadata$id[cntrl.inds]

#get only the cntrl cells from the counts matrix
cntrl.matrix <- counts[cntrl.ids]
```

```{r}
head(cntrl.matrix)
```


```{r}
mean <- rowSums(cntrl.matrix) / length(cntrl.matrix)
#ncol(cntrl.matrix) also gives the length # of columns
cntrl.means <- rowMeans(cntrl.matrix)
all( cntrl.means == mean )
names(cntrl.means) <- counts$ensgene 
cntrl.means[1:10]
```

## Lets do the same thing for the *treated* cells
```{r}
treated.inds = metadata$dex == "treated"

treated.ids <- metadata$id[treated.inds]

#get only the treated cells from the conts matrix
treated.matrix <- counts[treated.ids]
```
```{r}
treated.means <- rowMeans(treated.matrix)
names(treated.means) <- counts$ensgene
treated.means[1:10]
```
Combine the means into one df
```{r}
meanCounts <- data.frame(cntrl.means, treated.means)
colSums(meanCounts)
```


Now plot the means to see if there is any scatter, if there is that means the drug is having an effect on the RNA expression levels
```{r}
#bad plot, we should use a log plot!
plot(meanCounts)
```
```{r}
plot(meanCounts, log="xy")
```

To find what genes to look at lets see what the differences are in log means.
```{r}
#log2foldchange is used for its mathematical purposes, negative value means expression went down, positive means up.
meanCounts$log2fc <- log2(meanCounts[ ,"cntrl.means"]/meanCounts[ ,"treated.means"])
head(meanCounts)
```


```{r}
library(ggplot2)
```

```{r}
logplot <- ggplot(meanCounts, aes(x=cntrl.means, y = treated.means)) +geom_point()
logplot + scale_x_continuous(trans = "log2") + scale_y_continuous(trans = "log2")
```
Back to meanCounts. We should remove genes with NaN and inf means.

```{r}
toy <- meanCounts[1:6,]
which(toy[ ,1:2] == 0, arr.ind = TRUE )
#lets see if those values correspond correctly
```

```{r}
zero.vals <- which(meanCounts[ ,1:2] == 0, arr.ind = TRUE )
to.rm <- unique(zero.vals[,1])
myCounts <- meanCounts[-to.rm, ]
head(myCounts)
```

lets remove those 
```{r}
up.inds <- myCounts$log2fc > 2
down.inds <- myCounts$log2fc <(-2)
sum(up.inds)
sum(down.inds)
```
The amount of upregulated genes is `sum(up.inds)` and downregulated genes are `sum(down.inds)`

## Adding annotation data
```{r}
annotations <- read.csv("data/annotables_grch38.csv")
head(annotations)
```
### Use **merge()** to merge annotation data with *myCounts*
```{r}
myCounts.annotated <- merge(x=myCounts, y=annotations, by.x="row.names",by.y="ensgene")
head(myCounts.annotated)
```


## Using DESeq2
To do some proper analysis with statistical backing, p-values!
```{r}
library(DESeq2)
```

```{r}
#DESeq2 is driven by the DESeq2DataSet object
dds <- DESeqDataSetFromMatrix(countData = counts, colData = metadata, design = ~dex, tidy = TRUE)
```
```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

```{r}
summary(res)
```

Volcano plot helps show the genes that are signficantly changing
```{r}
#set all to gray
myColors <- rep("gray", nrow(res))
#set genes with l2fc greater than 2 to color red
myColors[abs(res$log2FoldChange)>2] <- "red"
#set the significant values to blue
colinds <- res$padj < 0.02 & abs(res$log2FoldChange) > 2
myColors[colinds] <- "blue"
#wnat the negativ love of p values, -log(res$padj) because p values are close together, negative so that the values are go in the positive Y direction
plot(res$log2FoldChange, -log(res$padj), col=myColors, ylab="-Log(P)", xlab="Log2FoldChange") 
```
lets use ggplot
```{r}
res$sig <- res$padj < 0.05 
ggplot(as.data.frame(res), aes(log2FoldChange, -1*log10(pvalue), col=sig)) + 
    geom_point() + ggtitle("Volcano plot")

#ggplot(as.data.frame(res) convert to dataframe 
#the values wanted to be used : aes(log2FoldChange, -1*log10(pvalue)) 
#color based on significance : col=sig)
```

```{r}

```

