---
title: "Pathway Analysis"
author: "Brandon Gonzalez"
date: "February 28, 2019"
output: html_document
---
# 	A microRNA network regulates expression and biosynthesis of CFTR and CFTR-ΔF508.
Ramachandran S, Karp PH, Jiang P, Ostedgaard LS et al. A microRNA network regulates expression and biosynthesis of wild-type and DeltaF508 mutant cystic fibrosis transmembrane conductance regulator. Proc Natl Acad Sci U S A 2012 Aug 14;109(33):13362-7. PMID: 22853952
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
#source
#biocLite("GEOquery")
#biiocLite("Biobase")
library(Biobase)
library(GEOquery)
library(limma)
library(DESeq2)
```

Open up the GEO file *even if it is compressed*
```{r}
gds4255 <- getGEO(filename='data/GDS4255_full.soft.gz', AnnotGPL = TRUE)
```
##Lets look at the GDS object from the GEO entry
There are two main things the GDS object gives us, meta data and a table for expression. From class 16 we recall that the DESeq2 workflow requires two tables, metaData table and the countData table, however expression data is different from count data..
```{r}
Meta(gds4255)
```
```{r}
Table(gds4255)[1:10, ]
```
```{r}
eset <- GDS2eSet(gds4255, getGPL= FALSE)

pData(eset)
```
gds4255

```{r}
Meta(gds4255)$sample_count
Meta(gds4255)$sample_organism
```

```{r}
gdsTable <- Table(gds4255)
colnames(gdsTable)
```

```{r}
row.names(gdsTable) <- gdsTable[,1]
head(gdsTable)
```

```{r}
eset <- GDS2eSet(gds4255, getGPL = FALSE)
malist <- GDS2MA(gds4255)
```
Lets see that we have the 12 expected samples
```{r}
sampleNames(eset)
```

```{r}
head(malist$M)
head(malist$A)
```

```{r}
ourData <- malist$M
row.names(ourData) <- gdsTable$ID_REF
head(ourData)
```



```{r}
pData(eset)[1:9,]
```

```{r}
design <- Columns(gds4255)[,1:3]
design <- design[1:9,]
design
```
```{r}
conDesign <- rep(0,length(design[,2]))
conDesign[design[,2] != "control"] = 1
conDesign
```



```{r}
dim(ourData)
subsetData <- ourData[,1:9]
dim(subsetData)
```

```{r}
fit <- lmFit(subsetData,conDesign)

```
```{r}
fit <- eBayes(fit)
```

```{r}
volcanoplot(fit)
```


```{r}
library(pathview)
library(gage)
library(gageData)
```


```{r}
genesOfInterest <- topTable(fit,number=100)
genesOfInterest$ID_REF <- row.names(genesOfInterest)

```
```{r}
genesOfInterest[1:20 , ]
```

```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```
```{r}
columns(org.Hs.eg.db)
```
```{r}
IdRefs <- row.names(genesOfInterest)
IdRefs
```
```{r}
geneTable <- gdsTable[IdRefs, ]
geneTable
```

Load up some pathways
```{r}
data("kegg.sets.hs")
```

## Assigning entrez IDs to the genesOfInterst.

lets take a peak at the original GEO data that we downloaded and see how the genes are IDd
```{r}
Table(gds4255)[1:20,]
```
It seems like geneBank ascenison is the only identifiers that is present for all the entries. Lets double check.

```{r}
Table(gds4255)$`GenBank Accession`
```


gage() takes a vector of named logfoldchanges and a set of pathways you want to compare over 

```{r}
gage(genesOfInterest$logFC

```

